Introduction to Markov state modeling with the PyEMMA software --- v\versionnumber

### 2018-09-13 manuscript/manuscript.tex
update to figure 2 (#136)

\begin{figure}
\includegraphics{figure_2}
\caption{Example analysis of the conformational dynamics of a pentapeptide backbone: (a)~The Trp-Leu-Ala-Leu-Leu pentapeptide in licorice representation~\cite{vmd}.
(b)~The VAMP-2 score indicates which of the tested featurizations contains the highest kinetic variance.
(c)~The sample density projected onto the first two time-lagged independent components (ICs) at lag time $\tau=0.5$~ns shows multiple density maxima and
(d)~the time series of the first two ICs of the first trajectory.}
\label{fig:io-to-tica}
\end{figure}

@@

In Markov state modeling our objective is to model the slow dynamics of a molecular process.
In order to approximate the slow dynamics in a statistically efficient manner,
a lower dimensional representation of our simulation data is necessary.
However, the features (e.g. torsion angles, distances or contacts) which best represent the slow dynamical modes of a given molecular system are unknown a priori~\cite{NoeClementiReview}.
Fortunately, the variational principle of conformational dynamics~\cite{noe-vac,nueske-vamk}
and the more general variational approach for Markov processes (VAMP)~\cite{vamp-preprint}
provide a systematic means to quantitatively compare multiple representations of the simulation data.
In particular, we can use a scalar score obtained using VAMP to directly compare the ability of certain features to capture slow dynamical modes in a particular molecular system.

@@

Subsequently, we perform TICA~\cite{tica,kinetic-maps} in order to reduce the dimension from the feature space,
which typically contains many degrees of freedom,
to a lower dimensional space that can be discretized with higher resolution and better statistical efficiency.
TICA is a special case of the variational principle~\cite{noe-vac,nueske-vamk} and is designed to find a projection preserving the long-timescale dynamics in the dataset.
Here, performing TICA on the backbone torsions at lag time~$0.5$~ns yields a four dimensional subspace using a~$95\%$ kinetic variance cutoff
(note that we perform a $\cos/\sin$-transformation of the torsions before TICA in order to preserve their periodicity).
The sample density projected onto the first two independent components (ICs) exhibits several maxima (Fig.~\ref{fig:io-to-tica}c).
Discrete jumps between the maxima can be observed by visualizing the transformation of the first trajectory into these ICs (Fig.~\ref{fig:io-to-tica}d).
We thus assume that our TICA-transformed backbone torsion features describe one or more metastable processes.

>>>>

\begin{figure}
\includegraphics{figure_2}
\caption{Example analysis of the conformational dynamics of a pentapeptide backbone: (a)~The Trp-Leu-Ala-Leu-Leu pentapeptide in licorice representation~\cite{vmd}.
(b)~The VAMP-2 score indicates which of the tested featurizations contains the highest kinetic variance.
(c)~The sample free energy projected onto the first two time-lagged independent components (ICs) at lag time $\tau=0.5$~ns shows multiple minima and
(d)~the time series of the first two ICs of the first trajectory show rare jumps.}
\label{fig:io-to-tica}
\end{figure}

@@

In Markov state modeling, our objective is to model the slow dynamics of a molecular process.
In order to approximate the slow dynamics in a statistically efficient manner,
a lower dimensional representation of our simulation data is necessary.
However, the features (e.g. torsion angles, distances or contacts) which best represent the slow dynamical modes of a given molecular system are unknown a priori~\cite{NoeClementiReview}.
Fortunately, the variational principle of conformational dynamics~\cite{noe-vac,nueske-vamk}
and the more general variational approach for Markov processes (VAMP)~\cite{vamp-preprint}
provide a systematic means to quantitatively compare multiple representations of the simulation data.
In particular, we can use a scalar score obtained using VAMP to directly compare the ability of certain features to capture slow dynamical modes in a particular molecular system.

@@

Subsequently, we perform TICA~\cite{tica,kinetic-maps} in order to reduce the dimension from the feature space,
which typically contains many degrees of freedom,
to a lower dimensional space that can be discretized with higher resolution and better statistical efficiency.
TICA is a special case of the variational principle~\cite{noe-vac,nueske-vamk} and is designed to find a projection preserving the long-timescale dynamics in the dataset.
Here, performing TICA on the backbone torsions at lag time~$0.5$~ns yields a four dimensional subspace using a~$95\%$ kinetic variance cutoff
(note that we perform a $\cos/\sin$-transformation of the torsions before TICA in order to preserve their periodicity).
The sample free energy projected onto the first two independent components (ICs) exhibits several minima (Fig.~\ref{fig:io-to-tica}c).
Discrete jumps between the minima can be observed by visualizing the transformation of the first trajectory into these ICs (Fig.~\ref{fig:io-to-tica}d).
We thus assume that our TICA-transformed backbone torsion features describe one or more metastable processes.
